*Supplementary Material*

### 2022-02-17 ms/manuscript.Rmd
Add PDF versions of ms and SI
For some weird reason need template .docx in two places :/

---
title: "An open and continuously updated fern tree of life (FTOL)"
output: 
  bookdown::word_document2:
    pandoc_args: [ "--csl", "frontiers.csl"]
    toc: no
    number_sections: no
    reference_docx: Frontiers_Template.docx
editor_options: 
  chunk_output_type: console
params:
  doc_type: manual
---

@@

We rooted the ML tree on bryophytes and estimated divergence dates using penalized likelihood with treePL [@Smith2012].
We selected `r n_calibrations_ferns` fern fossils to use as calibration points from a curated list of `r n_fern_fossils_total` fern fossil taxa.
In the case of redundant fossils (those calibrating the same node in the phylogeny; `r eg`, stems of families that are sister to each other, such as Dipteridaceae and Matoniaceae) we selected the fossil with the oldest age (`r ie`, the upper limit of the oldest stratigraphic period assigned to the fossil). 
We assigned fossils to lineages only after consulting the original publication (as opposed to only using the fossil species or genus) so we could reassess the identification relative to changes in taxonomy and hypotheses of phylogenetic relationships.
Taxon concepts may vary between the original publication and currently applied taxonomy (`r eg`, Dicksoniaceae *sensu lato* including other tree fern families used in description of the fossil but Dicksoniaceae *sensu stricto* used currently) and hypotheses of phylogenetic relationships among extant species may change (`r eg`, *Dennstaedtia* used in description of a fossil but this genus now known to be polyphyletic); in both cases, the resulting node to be calibrated for a given fossil will also change.
We also applied two calibration points outside of ferns including the most recent common ancestor (MRCA) of euphyllophytes (`r fossils$Euphyllophytes$age` Ma) and MRCA of tracheophytes (`r fossils$Tracheophytes$age` Ma), and fixed the root age (land plants) at 475 Ma [@Testo2016a].
All calibration points other than the root are minimum ages.

@@

# References
<!-- Insert <div id="refs"></div> to place the refs here instead of at the end -->
<div id="refs"></div>

# Figures

@@

>>>>

---
title: "An open and continuously updated fern tree of life (FTOL)"
mainfont: Roboto-Regular.ttf
fontsize: 11pt
mainfontoptions: 
  - BoldFont=Roboto-Bold.ttf
  - ItalicFont=Roboto-Italic.ttf
  - BoldItalicFont=Roboto-BoldItalic.ttf
output: 
  bookdown::pdf_document2:
    citation_package: default
    toc: no
    number_sections: yes
    fig_caption: yes
    keep_tex: no
    latex_engine: xelatex
    pandoc_args: [ "--csl", "frontiers.csl"]
  bookdown::word_document2:
    pandoc_args: [ "--csl", "frontiers.csl"]
    toc: no
    number_sections: no
    reference_docx: Frontiers_Template_SI.docx
header-includes: 
  - \usepackage{float} # keep figures from floating away
  - \makeatletter\renewcommand*{\fps@figure}{H}\makeatother # keep figures from floating away
  - \usepackage{fancyhdr} # enable header
  - \cfoot{\thepage} # footer
editor_options: 
  chunk_output_type: console
params:
  doc_type: manual
---

@@

We rooted the ML tree on bryophytes and estimated divergence dates using penalized likelihood with treePL [@Smith2012].
We selected `r n_calibrations_ferns` fern fossils to use as calibration points from a curated list of `r n_fern_fossils_total` fern fossil taxa (`r s_table("fossils")`).
In the case of redundant fossils (those calibrating the same node in the phylogeny; `r eg`, stems of families that are sister to each other, such as Dipteridaceae and Matoniaceae) we selected the fossil with the oldest age (`r ie`, the upper limit of the oldest stratigraphic period assigned to the fossil). 
We assigned fossils to lineages only after consulting the original publication (as opposed to only using the fossil species or genus) so we could reassess the identification relative to changes in taxonomy and hypotheses of phylogenetic relationships.
Taxon concepts may vary between the original publication and currently applied taxonomy (`r eg`, Dicksoniaceae *sensu lato* including other tree fern families used in description of the fossil but Dicksoniaceae *sensu stricto* used currently) and hypotheses of phylogenetic relationships among extant species may change (`r eg`, *Dennstaedtia* used in description of a fossil but this genus now known to be polyphyletic); in both cases, the resulting node to be calibrated for a given fossil will also change.
We also applied two calibration points outside of ferns including the most recent common ancestor (MRCA) of euphyllophytes (`r fossils$Euphyllophytes$age` Ma) and MRCA of tracheophytes (`r fossils$Tracheophytes$age` Ma), and fixed the root age (land plants) at 475 Ma [@Testo2016a].
All calibration points other than the root are minimum ages.

@@

# Acknowledgments {-}

Members of the Iwasaki lab provided helpful comments that improved the manuscript.
This study was supported by JSPS KAKENHI Grant Number 16H06279.
The authors thank Michael Hassler for maintaining the World Ferns taxonomic database and making it available to use for research.

# References {-}
<!-- Insert <div id="refs"></div> to place the refs here instead of at the end -->
<div id="refs"></div>

`r pagebreak_pdf(params)`

# Figures {-}

@@

`r pagebreak_pdf(params)`

### 2022-02-17 ms/si.Rmd
Add PDF versions of ms and SI
For some weird reason need template .docx in two places :/

---
title: "Supplemental Information"
mainfont: Roboto-Regular.ttf
fontsize: 11pt
mainfontoptions: 
  - BoldFont=Roboto-Bold.ttf
  - ItalicFont=Roboto-Italic.ttf
  - BoldItalicFont=Roboto-BoldItalic.ttf
output: 
  bookdown::pdf_document2:
    citation_package: default
    toc: no
    number_sections: no
    fig_caption: yes
    keep_tex: no
    latex_engine: xelatex
    pandoc_args: [ "--csl", "frontiers.csl"]
header-includes: 
  - \usepackage{float} # keep figures from floating away
  - \usepackage{booktabs} # need for kable tables
  - \usepackage{colortbl} # need for kable tables
  - \makeatletter\renewcommand*{\fps@figure}{H}\makeatother # keep figures from floating away
  - \usepackage{fancyhdr} # enable header
  - \pagestyle{fancy}  # enable header
  - \fancypagestyle{plain}{\pagestyle{fancy}} # show header on first page
  - \renewcommand{\headrulewidth}{0pt} # delete line beneath header
  - \fancyhead[RE,RO]{Nitta \textit{et al}. Fern Tree of Life SI} # add header on right side
  - \fancyhead[LO,LE]{} # suppress automatic section header on left side
  - \cfoot{\thepage} # footer
editor_options: 
  chunk_output_type: console
params:
  doc_type: manual
---

@@

```{r setup-si, include = FALSE}
knitr::opts_chunk$set(
	echo = FALSE, warning = FALSE, message = FALSE, results = 'hide')
# Load packages
library(ggupset)
library(kableExtra)
source(here::here("R/ms_packages.R"))
```

@@

tar_load(
	c(
		sanger_alignment_tbl,
		sanger_sampling,
		sanger_accessions_selection,
		plastome_tree,
		plastid_tree_dated),
	store = ftol_cache
)

# Targets from MS workflow
tar_load(
	c(
		gb_species_by_year,
		ref_files,
		other_dates,
		du_dates_all,
		family_stem_ages
	),
	store = here::here("_targets")
)
```

@@

```{r join_by, results = "as-is"}
# Make table showing method of joining accessions across loci
sanger_accessions_selection %>% 
	left_join(select(sanger_sampling, species, outgroup), by = "species") %>%
	# FIXME why is this NA?
	# filter(is.na(outgroup)) %>%
	filter(outgroup == FALSE) %>%
	select(-outgroup) %>%
	assert(not_na, join_by, species) %>%
	assert(is_uniq, species) %>%
	tabyl(join_by) %>%
	arrange(n) %>%
	adorn_totals() %>%
	mutate(
		join_by = str_to_sentence(join_by) %>%
			str_replace_all("Monophy", "Monophyletic"),
		n = number(n),
		percent = case_when(
			join_by != "Total" ~ percent(percent) %>% 
				str_replace_all("%", "\\\\%"),
			TRUE ~ ""
		)
	) %>%
	rename(
		"Method" = join_by,
		`\\textit{n}` = n,
		`\\%` = percent
	) %>%
  kbl(booktabs = TRUE, escape = FALSE, format = "latex") %>%
  kable_styling(latex_options = c("striped", "HOLD_position"))
```

@@

\clearpage

@@

>>>>

---
title: "*Supplementary Material*"
mainfont: Roboto-Regular.ttf
fontsize: 11pt
mainfontoptions: 
  - BoldFont=Roboto-Bold.ttf
  - ItalicFont=Roboto-Italic.ttf
  - BoldItalicFont=Roboto-BoldItalic.ttf
output: 
  bookdown::pdf_document2:
    citation_package: default
    toc: no
    number_sections: no
    fig_caption: yes
    keep_tex: no
    latex_engine: xelatex
    pandoc_args: [ "--csl", "frontiers.csl"]
  bookdown::word_document2:
    pandoc_args: [ "--csl", "frontiers.csl"]
    toc: no
    number_sections: no
    reference_docx: Frontiers_Template_SI.docx
header-includes: 
  - \usepackage{float} # keep figures from floating away
  - \usepackage{booktabs} # need for kable tables
  - \usepackage{colortbl} # need for kable tables
  - \makeatletter\renewcommand*{\fps@figure}{H}\makeatother # keep figures from floating away
  - \usepackage{fancyhdr} # enable header
  - \pagestyle{fancy}  # enable header
  - \fancypagestyle{plain}{\pagestyle{fancy}} # show header on first page
  - \renewcommand{\headrulewidth}{0pt} # delete line beneath header
  - \fancyhead[RE,RO]{Supplementary Material} # add header on right side
  - \fancyhead[LO,LE]{} # suppress automatic section header on left side
  - \cfoot{\thepage} # footer
editor_options: 
  chunk_output_type: console
params:
  doc_type: manual
---

@@

```{r setup-si, include = FALSE}
knitr::opts_chunk$set(
	echo = FALSE, warning = FALSE, message = FALSE, results = 'hide')
# Load packages
library(ggupset)
library(flextable)
source(here::here("R/ms_packages.R"))
```

@@

tar_load(
	c(
		sanger_alignment_tbl,
		sanger_sampling,
		sanger_accessions_selection,
		plastome_tree,
		plastid_tree_dated,
		fossil_calibration_tips),
	store = ftol_cache
)

# Targets from MS workflow
tar_load(
	c(
		gb_species_by_year,
		ref_files,
		other_dates,
		du_dates_all,
		family_stem_ages,
		fossil_ferns_raw
	),
	store = here::here("_targets")
)
```

@@

```{r fossils}
# Filter raw fossil data to just fossils used, write out as CSV
fossils_table <-
fossil_ferns_raw %>%
	inner_join(
		select(fossil_calibration_tips, fossil_taxon), by = "fossil_taxon") %>%
	verify(nrow(.) == nrow(fossil_calibration_tips)) %>%
	rowwise() %>%
	mutate(
		note = jntools::paste3(notes_21, notes_22sep = ". ") %>%
			str_replace_all("\\.\\.", "\\.") %>%
			na_if(".") %>%
			str_replace_all(" \\.$", "\\.")
	) %>%
	ungroup() %>%
	select(
		n_fos, minimum_age, node_calibrated, fossil_taxon, 
		lower = lower_limit_of_oldest_stratigraphic_age,
		upper = upper_limit_of_oldest_stratigraphic_age,
		stratigraphic_age, 
		country, type_locality, formation,
		organs, specimen_holotype, collection,
		reference, note
	)

write_csv(fossils_table, result_file(s_table("fossils"), "csv"))
```

**`r s_table("fossils")`**. Fossils used as calibration points for molecular dating (`r result_file(s_table("fossils"), "csv")`). "n_fos" is the fossil ID number used in our fossil database (https://github.com/pteridophy/fossils). "minimum_age" (Ma) is the age used for calibration; "lower" is the lower limit of the oldest stratigraphic age for the fossil; "upper" is the upper limit of the oldest stratigraphic age for the fossil.

\newpage

```{r join-by, results = "markup"}
# Make table showing method of joining accessions across loci
sanger_accessions_selection %>% 
	left_join(select(sanger_sampling, species, outgroup), by = "species") %>%
	# FIXME why is this NA?
	# filter(is.na(outgroup)) %>%
	filter(outgroup == FALSE) %>%
	select(-outgroup) %>%
	assert(not_na, join_by, species) %>%
	assert(is_uniq, species) %>%
	tabyl(join_by) %>%
	arrange(n) %>%
	adorn_totals() %>%
	mutate(
		join_by = str_to_sentence(join_by) %>%
			str_replace_all("Monophy", "Monophyletic"),
		n = number(n),
		percent = case_when(
			join_by != "Total" ~ percent(percent),
			TRUE ~ ""
		)
	) %>%
	rename(
		"Method" = join_by,
		`%` = percent) %>%
	flextable() %>%
	italic(1,2, part = "header") %>%
	autofit()
```

@@

\newpage

@@

\newpage
